---
output:
  html_document: default
  pdf_document: default
---
Red Wine Data Analysis by Monica London
========================================================

The purpose of this analysis is to explore a dataset featuring characteristics 
about red wine.

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
library(knitr)
library(ggplot2)
library(gridExtra)
library(grid)
library(dplyr)
library(psych)
library(ggcorrplot)
library(RColorBrewer)
setwd('~/Downloads')

```

```{r echo=FALSE, message=FALSE, warning=FALSE, Load_the_Data}

RedWine <- read.csv('wineQualityReds.csv')

```

# Summary

```{r echo=FALSE, message=FALSE, warning=FALSE, Summary}

summary(RedWine)

```

#Structure
```{r}
str(RedWine)
```

The red wine data set contains nearly 1600 observations of 13 variables.

```{r echo=FALSE, message=FALSE, warning=FALSE, Delete_Variable}

RedWine <- select(RedWine, -X)
```

#Univariate Plots Section

```{r echo=FALSE, message=FALSE, warning=FALSE, Univariate_Plots1}
ggplot(aes(x = quality), data = RedWine) +
  geom_histogram(binwidth = .5) +
  scale_x_continuous(breaks = seq(0, 10, 1))

```

We can see the distribution of quality ratings has a minimum of 3 and a maximum 
of 8, with most ratings at 5 or 6. Surprisingly, there are no ratings of 1, 2, 
9, or 10. I would have expected a larger range of quality ratings with such a 
large data set.

```{r echo=FALSE, message=FALSE, warning=FALSE, Univariate_Plots2}
RedWine$quality.level <- cut(RedWine$quality, breaks = c(0, 4, 6, 10))
qplot(data = RedWine, x = quality.level)

```

I divided the data by quality level, with 0-3, 4-6, and 7-10 being the three
levels. We can see that the vast majority of observations fall in the medium 
quality level.

```{r echo=FALSE, message=FALSE, warning=FALSE, Univariate_Plots3}
fixed_acidity = qplot(data = RedWine, x = fixed.acidity, xlab = "Fixed Acidity", 
                      fill = I("red"), col = I("black"))
volatile_acidity = qplot(data = RedWine, x = volatile.acidity, 
                        xlab = "Volatile Acidity", fill = I("red"), 
                        col = I("black"))
citric_acid = qplot(data = RedWine, x = citric.acid, xlab = "Citric Acid", 
                    fill = I("red"), col = I("black"))
residual_sugar = qplot(data = RedWine, x = residual.sugar, binwidth = .5, 
                       xlab = "Residual Sugar", fill = I("red"), 
                       col = I("black"))
chlorides = qplot(data = RedWine, x = chlorides, xlab = "Chlorides", 
                  fill = I("red"), col = I("black"))
alcohol = qplot(data = RedWine, x = alcohol, binwidth = .25, xlab = "Alcohol", 
                fill = I("red"), col = I("black"))
density = qplot(data = RedWine, x = density, xlab = "Density", fill = I("red"), 
                col = I("black"))
pH = qplot(data = RedWine, x = pH, binwidth = .1, xlab = "pH", fill = I("red"), 
            col = I("black"))
sulphates = qplot(data = RedWine, x = sulphates, binwidth = .05, 
                  xlab = "Sulphates", fill = I("red"), col = I("black"))
free_sulfur_dioxide = qplot(data = RedWine, x = free.sulfur.dioxide, 
                            binwidth = 2, xlab = "Free Sulfur Dioxide", 
                            fill = I("red"), col = I("black"))
total_sulfur_dioxide = qplot(data = RedWine, x = total.sulfur.dioxide, 
                             xlab = "Total Sulfur Dioxide", fill = I("red"), 
                             col = I("black"))

grid.arrange(fixed_acidity, volatile_acidity, citric_acid, residual_sugar, 
             chlorides, alcohol, density, pH, sulphates, free_sulfur_dioxide, 
             total_sulfur_dioxide, ncol=3)
```

We can see that the Density and pH plots are the most normally distributed. 
Thee majority of pH levels fall between 3.0 - 3.5. Many of the plots are skewed 
to the right, including Free Sulfur Dioxide, Total Sulfur Dioxide. The majority 
of wines havie less than 100 in total sulfur dioxide. Several of the plots are 
long tailed, such as Residual Sugar and Chlorides.


```{r echo=FALSE, message=FALSE, warning=FALSE, Univariate_Plots4} 
log10_residual_sugar = qplot(data = RedWine, x = log10(residual.sugar), 
                             binwidth = .025, xlab = "Residual Sugar (log10)", 
                             fill = I("blue"), col = I("black"))
log10_chlorides = qplot(data = RedWine, x = log10(chlorides), 
                        xlab = "Chlorides (log10)", fill = I("blue"), col = 
                          I("black"))
log10_f_sulfur_dioxide = qplot(data = RedWine, x = log10(free.sulfur.dioxide), 
                               binwidth = .1, 
                               xlab = "Free Sulfur Dioxide (log10)", 
                               fill = I("blue"), col = I("black"))
log10_t_sulfur_dioxide = qplot(data = RedWine, x = log10(total.sulfur.dioxide), 
                               binwidth = .05, 
                               xlab = "Total Sulfur Dioxide (log10)", 
                               fill = I("blue"), col = I("black"))

grid.arrange(residual_sugar, log10_residual_sugar, chlorides, log10_chlorides, 
             free_sulfur_dioxide, log10_f_sulfur_dioxide, total_sulfur_dioxide, 
             log10_t_sulfur_dioxide, ncol=2)
```

The above plots compare the plots before and after transformation. The data for 
residual sugar, chlorides, free sulfur dioxide, and total sulfur dioxide becomes 
more normally distributed after applying log10.

# Univariate Analysis

### What is the structure of your dataset?

There are 1599 wines in the dataset with 12 features (fixed acidity, volatile 
acidity, citric acid, residual sugar, chlorides, free sulfur dioxide, total 
sulfur dioxide, density, pH, sulfates, alcohol, and quality). All of the 
variables are numeric with the exception of quality, which is in integer form.

####Other observations:

Most of the wines have a quality of 5 or 6.

The 3rd quartile of residual sugar levels is 2.6, although there are a few major 
outliers, with the maximum residual sugar level of 15.5. I'm interested to see 
if higher residual sugar wines tend to have lower or higher quality.

Most wines have an alcohol content of less than 12%. This surprises me, given 
that the majority of red wines I'm familiar with have alcohol contents above 
13.5%.

Many of the wines have 0 citric acid.

### What is/are the main feature(s) of interest in your dataset?

The main feature of interest in the dataset are quality, and I'd like to 
determine which variables impact quality ratings the most. I suspect alcohol, 
residual sugar, and pH contribute to quality ratings, as they seem to be 
features you may be able to decipher during wine tastings.

### What other features in the dataset do you think will help support your \
investigation into your feature(s) of interest?

From research into what contributes to the taste of wine, I discovered that 
sweetness, acidity, tannin, alcohol, and body are the main features. 
In addition to pH, I think fixed acidity and volatile acidity may contribute to 
the acidity of wine.

### Did you create any new variables from existing variables in the dataset?

Yes, I created a new variable called quality level, which cuts the quality 
levels into low (3, 4), medium (5, 6), and high (7,8).

### Of the features you investigated, were there any unusual distributions? \
Did you perform any operations on the data to tidy, adjust, or change the form \
of the data? If so, why did you do this?

I deleted column X because it was simply a repeat of the index.

I applied log10 to residual sugar, chlorides, free sulfur dioxide, and total 
sulfur dioxide in order to normalize the distributions.

# Bivariate Plots Section

I analyzed the following bivariate relationships:

Quality vs. Alcohol
Quality vs. pH
Quality vs. Residual Sugar
Quality vs. Fixed Acidity
Quality vs. Volatile Acidity
Residual Sugar vs. Alcohol
Residual Sugar vs. pH
pH vs. Alcohol
Fixed Acidity vs. Density
Fixed Acidity vs pH
pH vs. Citric Acid
Quality Level vs. Alcohol
Quality Level vs. pH
Quality Level vs. Residual Sugar

```{r echo=FALSE, message=FALSE, warning=FALSE, Correlogram}
corr = cor(subset(RedWine, select = -quality.level))
ggcorrplot(corr, hc.order = TRUE, 
           type = "lower", 
           lab = TRUE, 
           lab_size = 3, 
           method="circle", 
           colors = c("tomato2", "white", "springgreen3"), 
           title="Correlogram of Red Wine", 
           ggtheme=theme_bw)
```

The correlogram indicates that the majority of relationships between variables 
are not highly correlated. The strongest relationships appear to be density vs. 
fixed acidity (r^2 = 0.67), citric acid vs. fixed acidity (r^2 = 0.67, pH vs. 
fixed acidity (r^2 = -0.68), and total sulfur dioxide vs. free sulfur dioxide 
(r^2 = 0.67). A correlation between citric acidy and fixed acidity is not 
surprising as they are both acids. Free sulfur dioxide is part of total sulfur 
dioxide so a correlation is expected. pH is a measure of acidity so the 
correlation between pH and fixed acidity is not surprising either. I am unsure 
what would cause a correlation between density and fixed acidity, but it could 
be that higher acidic liquid is more dense than lower acidic liquid.

```{r echo=FALSE, message=FALSE, warning=FALSE, Bivariate_Plots1}
ggplot(aes(x = alcohol, y = quality), data = RedWine) + 
  geom_point(alpha = .5, position = 'jitter') + 
  labs(title = "Quality vs. Alcohol") + 
  theme(plot.title = element_text(hjust = 0.5))
```

```{r echo=FALSE, message=FALSE, warning=FALSE, Bivariate_Plots2}
ggplot(aes(x = pH, y = quality), data = RedWine) + 
  geom_point(alpha = .5, position = 'jitter') + 
  labs(title = "Quality vs. pH") + theme(plot.title = element_text(hjust = 0.5))
```

```{r echo=FALSE, message=FALSE, warning=FALSE, Bivariate_Plots3}
ggplot(aes(x = log10(residual.sugar), y = quality), data = RedWine) + 
  geom_point(alpha = .5, position = 'jitter') + 
  labs(title = "Log10(Residual Sugar) vs. Quality") + 
  theme(plot.title = element_text(hjust = 0.5))
```

```{r echo=FALSE, message=FALSE, warning=FALSE, Bivariate_Plots4}
ggplot(aes(x = fixed.acidity, y = quality), data = RedWine) + 
  geom_point(alpha = .5, position = 'jitter') + 
  labs(title = "Quality vs. Fixed Acidity") + 
  theme(plot.title = element_text(hjust = 0.5))
```

```{r echo=FALSE, message=FALSE, warning=FALSE, Bivariate_Plots5}
ggplot(aes(x = volatile.acidity, y = quality), data = RedWine) + 
  geom_point(alpha = .3, position = 'jitter') + 
  labs(title = "Quality vs. Volatile Acidity") + 
  theme(plot.title = element_text(hjust = 0.5))
```

```{r echo=FALSE, message=FALSE, warning=FALSE, Bivariate_Plots6}
ggplot(aes(x = log10(residual.sugar), y = alcohol), data = RedWine) + 
  geom_point(alpha = .5, position = 'jitter') + 
  labs(title = "Log10(Residual Sugar) vs. Alcohol") + 
  theme(plot.title = element_text(hjust = 0.5))
```


```{r echo=FALSE, message=FALSE, warning=FALSE, Bivariate_Plots7}
ggplot(aes(x = pH, y = log10(residual.sugar), binwidth = .0005), data = RedWine) + 
  geom_point(alpha = .5, size = .8, position = 'jitter') + 
  labs(title = "Log10(Residual Sugar) vs. pH") + 
  theme(plot.title = element_text(hjust = 0.5))
```


```{r echo=FALSE, message=FALSE, warning=FALSE, Bivariate_Plots8}
ggplot(aes(x = alcohol, y = pH), data = RedWine) + 
  geom_point(alpha = 1/2, position = 'jitter') + 
  labs(title = "pH vs. Alcohol") + theme(plot.title = element_text(hjust = 0.5))
```


```{r echo=FALSE, message=FALSE, warning=FALSE, Bivariate_Plots9}
ggplot(aes(x = fixed.acidity, y = density), data = RedWine) + 
  geom_point(alpha = .5, position = 'jitter') + geom_smooth() + 
  labs(title = "Fixed Acidity vs. Density") + 
  theme(plot.title = element_text(hjust = 0.5))
```

```{r echo=FALSE, message=FALSE, warning=FALSE, Bivariate_Plots10}
ggplot(aes(x = fixed.acidity, y = pH), data = RedWine) + 
  geom_point(alpha = .5, position = 'jitter') + geom_smooth() + 
  labs(title = "Fixed Acidity vs. pH") + 
  theme(plot.title = element_text(hjust = 0.5))
```

```{r echo=FALSE, message=FALSE, warning=FALSE, Bivariate_Plots11}
qplot(x = quality.level, y = alcohol, data = RedWine) + 
  geom_boxplot(outlier.colour = 'red') + 
  labs(title = "Quality Level vs. Alcohol") + 
  theme(plot.title = element_text(hjust = 0.5))
```

It is interesting to see how the alcohol level tends to be much higher in the 
higher quality wines than the medium or low quality wines.

```{r echo=FALSE, message=FALSE, warning=FALSE, Bivariate_Plots12}
qplot(x = quality.level, y = pH, data = RedWine) + 
  geom_boxplot(outlier.colour = 'red') + labs(title = "Quality Level vs. pH") + 
  theme(plot.title = element_text(hjust = 0.5))
```

The median pH level decreases as the quality increases. The data is also more 
compact at the higheset quality level. As the quality level increases, the pH 
range decreases.

```{r echo=FALSE, message=FALSE, warning=FALSE, Bivariate_Plots13}
qplot(x = quality.level, y = log10(residual.sugar), data = RedWine) + 
  geom_boxplot(outlier.colour = 'red') + 
  labs(title = "Quality Level vs. Residual Sugar") + 
  theme(plot.title = element_text(hjust = 0.5))
```

The range of outliers (plotted in red) is large in this plot, especially in the 
medium quality level. All of the outliers in all quality levels are high 
outliers; they have very high levels of residual sugar rather than very low 
levels of residual sugar.

```{r echo=FALSE, message=FALSE, warning=FALSE, Bivariate_Plots14}
qplot(x = quality.level, y = volatile.acidity, data = RedWine) + 
  geom_boxplot(outlier.colour = 'red') + 
  labs(title = "Quality Level vs. Volatile Acidity") + 
  theme(plot.title = element_text(hjust = 0.5))
```

Both the IQR and median volatile acidity decreases as quality level increases.

# Bivariate Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. How did the feature(s) of interest vary with other features in \
the dataset?

The main feature of interest in this analysis is quality, and if any features 
show an affect on quality. The correlogram shows that the highest r^2 value
between quality and any other feature is alcohol (r^2 = 0.48). While you could
see this trend in alcohol vs. quality scatterplot, the extreme number of
quality ratings of 5 and 6 made it difficult to see the true relationship
between alcohol and quality in the higher quality wines. The he quality level
vs. alcohol boxplot shows this relationship much better, with a clear increase
in median alcohol levels in the highest quality wines.

The relationship between quality vs. pH has an r^2 value of 0.06, which
indicates practically zero correlation, and the corresponding scatterplot
confirms this. However, when pH is compared to quality levels, there is a
pattern in the boxplot. The median pH levels seem to decrease as the quality
level increases, especially between the lower quality and medium quality wines.

The correlogram indicates that there is no correlation (r^2 = 0.01) between
residual sugar and quality. Even after transforming the residual sugar data
using log10 and plotting it against quality levels, there seems to be no clear
relationship betwen residual sugar and quality. 

One of the most poignant bivariate relationships discovered was the relationship 
between quality level and volatile acidity. The correlogram shows an r^2 value 
of -0.39 between quality and volatile acidity. The scatterplot shows this 
moderate correlation, but the relationship is much clearer when the data is 
grouped by quality level in the volatile acidity vs. quality level boxplot.

### Did you observe any interesting relationships between the other features \
(not the main feature(s) of interest)?

Some of the most interesting relationships were between variables that were not
the main feature of interest. In fact, three of the four strongest r^2 values 
included fixed acidity vs. another variable. Fixed acidity had the strongest 
correlations with density, citric acid, and pH. As discussed previously, this
is not surprising given that many of the variables are either acid themselves or
a measure of acidity.

### What was the strongest relationship you found?

The strongest relationship, according to the r^2 value, is between pH and 
fixed acidity (r^2 = 0.68). However, once the data was cut into quality levels, 
the plots indicate that there are strong relationships between quality level and 
alcohol, quality level and volatile acidity, and quality level and pH.

# Multivariate Plots Section

```{r echo=FALSE, message=FALSE, warning=FALSE, Multivariate_Plots1}
ggplot(aes(x = fixed.acidity, y = density, colour = quality.level), data = RedWine) +
  geom_point(alpha = .5, position = 'jitter') +
  scale_color_brewer(type = 'seq', palette = 'Dark2',
    guide = guide_legend(title = 'Quality Level', reverse = T,
    override.aes = list(alpha = 1, size = 2))) +
  ggtitle('Density by Fixed Acidity and Quality Level') + 
  theme(plot.title = element_text(hjust = 0.5)) + geom_smooth()
```

Because the majority of the data has a medium quality level, the data is highly 
clustered. The smoother shows a slightly higher fixed acidity vs. density ratio 
for higher quality level wines vs. medium or lower quality level wines.

```{r echo=FALSE, message=FALSE, warning=FALSE, Multivariate_Plots2}
ggplot(aes(x = alcohol, y = pH, colour = quality.level), data = RedWine) +
  geom_point(alpha = .5, size = 1.5, position = 'jitter') + ylim(3, 4) +
  scale_color_brewer(type = 'qual', palette = 'Dark2',
    guide = guide_legend(title = 'Quality Level', reverse = T)) + 
  ggtitle('Alcohol by pH and Quality Level') + 
  theme(plot.title = element_text(hjust = 0.5))
    
  
```

This plot doesn't show strong trends, but it does show how the majority of the
data falls in the lower alcohol, lower pH quadrant of the chart.

```{r echo=FALSE, message=FALSE, warning=FALSE, Multivariate_Plots3}
ggplot(aes(x = pH, y = citric.acid, colour = quality.level), data = RedWine) +
  geom_point(alpha = .5) + xlim(3, 3.75) +
  scale_color_brewer(type = 'qual', palette = 'Dark2',
    guide = guide_legend(title = 'Quality Level', reverse = T)) +
  ggtitle('Citric Acid by pH and Quality Level') + 
  theme(plot.title = element_text(hjust = 0.5)) + geom_smooth()
```

This plot shows some differences in the relationship between pH and citric acid
by quality level. 

```{r echo=FALSE, message=FALSE, warning=FALSE, Multivariate_Plots4}
ggplot(aes(x = fixed.acidity, y = pH, colour = quality.level), data = RedWine) +
  geom_point(alpha = .5, position = 'jitter') +
  scale_color_brewer(type = 'seq', palette = 'Dark2',
    guide = guide_legend(title = 'Quality Level', reverse = T,
    override.aes = list(alpha = 1, size = 2))) +
  ggtitle('pH by Fixed Acidity and Quality Level') + 
  theme(plot.title = element_text(hjust = 0.5))
```

The relationship between pH and fixed acidity seems to be uniform across all
quality levels.

```{r echo=FALSE, message=FALSE, warning=FALSE, Model}
library(memisc)

	m1<-lm(quality ~ alcohol, data = RedWine)
    m2 <- update(m1, ~ . + pH)
    m3 <- update(m2, ~ . + I(log10(residual.sugar)))
    m4 <- update(m3, ~ . + residual.sugar)
    m5 <- update(m4, ~ . + fixed.acidity)
    m6 <- update(m5, ~ . + volatile.acidity)
    m7 <- update(m6, ~ . + I(log10(total.sulfur.dioxide)))
    m8 <- update(m7, ~ . + total.sulfur.dioxide)
    m9 <- update(m8, ~ . + citric.acid)
    m10 <- update(m9, ~ . + I(log10(chlorides)))
    m11 <- update(m10, ~ . + I(log10(free.sulfur.dioxide)))
    m12 <- update(m11, ~ . + sulphates)
    mtable(m1, m2, m3, m4, m5, m6, m7, m8, m9, m10, m11, m12)

```

# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. Were there features that strengthened each other in terms of \
looking at your feature(s) of interest?

When looking at fixed acidity vs. citric acid in terms of quality levels, there
does seem to be a relationship. For any given pH, the citric acid level appears 
to increase as the quality level increases. This relationship isn't as clear as 
the pH levelincreases above 3.5. This may be because of fewer data points at 
that level.

In the density vs. fixed acidity in terms of quality level plot, there is a very
relationship. For a given fixed acidity level below 12, the average density 
level is higher for lower quality wines than higher quality wines.

### Were there any interesting or surprising interactions between features?

In the density vs. fixed acidity in terms of quality level plot, the smoother
for the lowest quality wines does not appear linear. It almost appears 
logarithmic, rising in density value slower as the fixed acidity value 
increases.

### Did you create any models with your dataset? Discuss the \
strengths and limitations of your model.

I created a model to predict quality with several variables, including alcohol, 
pH, residual sugar, fixed acidity, volatile acidity, total sulfur dioxide, 
citric acid, chlorides, free sulfur dioxide, and sulphates. The r^2 value of
the model is 0.361. Because quality ratings are chosen by humans and are not
scientic, an r^2 value of 0.361 is relatively strong. 

The number of observations with 3, 4, 7, and 8 quality ratings is so much lower 
than the number of observations with 5 and 6 quality ratings. A bigger overall
data set and more observations in the lower and higher quality ratings would
improve the model.

# Final Plots and Summary

### Plot One
```{r echo=FALSE, message=FALSE, warning=FALSE, Plot_One}
qplot(x = quality.level, y = alcohol, data = RedWine) + 
  geom_boxplot(outlier.colour = 'red') + 
  labs(x = "Quality Level", y = "Alcohol (%)", 
       title = "Quality Level vs. Alcohol") +
  theme(plot.title = element_text(hjust = 0.5))
```

### Description One
This plot shows a similar median alcohol level for both low and medium quality 
wines. Interestingly the median alcohol level spikes much higher for the highest 
quality wines.

### Plot Two
```{r echo=FALSE, message=FALSE, warning=FALSE, Plot_Two}
qplot(x = quality.level, y = pH, data = RedWine) + 
  geom_boxplot(outlier.colour = 'red') + 
  labs(x = "Quality Level", title = "Quality Level vs. pH Level") + 
  theme(plot.title = element_text(hjust = 0.5))
```

### Description Two
This plot indicates that the median pH level steadily decreases as the quality 
level increases.

### Plot Three
```{r echo=FALSE, message=FALSE, warning=FALSE, Plot_Three}
ggplot(aes(x = fixed.acidity, y = pH), data = RedWine) + 
  geom_point(alpha = 1/2, position = 'jitter') + geom_smooth() + 
  labs(x = "Fixed Acidity (g/dm^3)", title = "Fixed Acidity vs. pH") + 
  theme(plot.title = element_text(hjust = 0.5))
```

### Description Three
This plot is notable because it visualizes the relationship between the two 
variables with the strongest correlation in the data set. Fixed acidity and pH 
have an r^2 value of -0.68. This relationship makes a lot of sense because a pH 
level is a measure of acidity; a lower pH indicates a substance is more acidic 
and a higher pH indicates a substance is more basic. This plot confirms this 
relationship.
------

# Reflection

I wanted to find out which variables most impacted quality. There were several 
insights I found during the exploration of this data set.

Alcohol and Quality: There is a clear relationship between quality level and 
alcohol content, but only for the highest quality wines. This relationship is
unclear until quality is separated into quality levels. 

pH and Quality: There is a negative relationship between pH and quality levels. 
This is unclear until quality is separated into quality levels.

Volatile Acidity and Quality: There is a very strong relationship between 
volatile relationship and quality levels. The relationship is only moderate 
when quality is not divided into levels. 

The correlogram was very helpful in showing correlations between variables 
except for quality. In several cases, the relationship between quality and a 
specific variable was unclear until quality was separated into quality levels. 

It would be most helpful to know the type of red wine, such as Cabernet 
Sauvignon, Pinot Noir, Merlot, etc. It is very difficult to analyze trends when
the type of the wine is unknown. For example, a certain wine type may be 
expected to have more alcohol and therefore someone rating the quality of that 
wine would rate it more favorably than someone rating a quality of wine that was 
expected to have a lower alcohol content. Furthermore, it would be interesting 
to analyze wines from different parts of the world to see if there is a 
relationship between quality or any of the variables and region.
